---
title: "Reproducible Research: Peer Assessment 1"
author: "Centrallimit"
output: html_document
keep_md: true
theme: "spacelab"
---


## Loading and preprocessing the data

This document was created on `r weekdays(Sys.Date())`, `r Sys.Date()`.

This section describes the data integration process. 

```{r echo = TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

unzip("./activity.zip") # contains activity.csv
df_act <- tbl_df(read.csv("activity.csv", header = TRUE, sep = ","))
mutate(df_act, date=as.Date(date))

```


## What is mean total number of steps taken per day?


```{r echo = TRUE, message=FALSE, warning=FALSE}
steps_sum <- df_act %>% 
        group_by(date) %>%
        summarise(stepssum=sum(steps,na.rm=TRUE))
c_mean <- round(mean(steps_sum$stepssum, na.rm = TRUE), 2)
c_med <- round(median(steps_sum$stepssum, na.rm = TRUE), 2)
```
The mean total number of steps taken per day is `r c_mean` and the median is `r c_med`. 

```{r echo = TRUE, message=FALSE, warning=FALSE}
ggplot(steps_sum, aes(x=stepssum)) +
geom_histogram(colour="black", fill="red")
    

```

## What is the average daily activity pattern?

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Grouping and computation of the mean 
steps_day <-  df_act %>%
        group_by(interval) %>%
        summarise(mean_int=mean(steps, na.rm = TRUE))

# Identify the interval with the maximum value of steps
c_max <- steps_day[which.max(steps_day$mean_int), ]    
# Time series plot of daily average 
ggplot(steps_day, aes(interval, mean_int)) + geom_line(color = "red") + xlab("Interval") + ylab("Mean")

```

The interval, that on average across all the days in the dataset, contains the maximum number of steps is `r c_max[1,1]` with an average of `r c_max[1,2]` steps.

## Imputing missing values
First of all, we have to compute the number of missing values. 

```{r echo = TRUE, warning=FALSE, message=FALSE}
# Load library that contains the impute() function
library(Hmisc)

# Number of datasets that have missing values
count_c <- sum(!complete.cases(df_act))
# Imputing missing values of steps with the mean value
df_act_i <- df_act %>%
        mutate(steps = impute(df_act$steps, mean))

steps_sum_i <- df_act_i %>% 
        group_by(date) %>%
        summarise(stepssum_i=sum(steps,na.rm=TRUE))


c_mean_i <- round(mean(steps_sum_i$stepssum_i, na.rm = TRUE), 2)

c_med_i <- round(median(steps_sum_i$stepssum_i, na.rm = TRUE), 2)

```
       
The are `r count_c` datasets with missing values in the sample. I used the mean to impute 
the missing values. The new mean and and median after the imputation process are `r round(c_mean_i,2)`
and `r round(c_med_i,2)`.
The new histogram looks like this.

```{r echo = TRUE, message=FALSE, warning=FALSE}
ggplot(steps_sum_i, aes(x=stepssum_i)) +
geom_histogram(colour="black", fill="blue")
    

```



## Are there differences in activity patterns between weekdays and weekends?

```{r echo = TRUE, warning=FALSE, message=FALSE}
# Vector with weekdays in German language
# ATTENTION: You may have to use the English version to reproduce the results
weekday <- c('Montag', 'Dienstag', 'Mittwoch', 'Donnerstag', 'Freitag')
# weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

df_last <- df_act_i %>%
        mutate(wDay = factor((weekdays(as.Date(df_act_i$date)) %in% weekday),
                               levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')))

# Grouping and computation of the mean 
steps_day_i <-  df_last %>%
        group_by(interval, wDay) %>%
        summarise(mean_int_i=mean(steps, na.rm = TRUE))


```


```{r echo=TRUE}
# Time series plot of daily average 
ggplot(steps_day_i, aes(interval, mean_int_i)) + 
        geom_line(color = "blue") + 
        facet_grid(wDay ~ .) + xlab("Interval") + ylab("Mean")

```

