---
title: "Reproducible Research: Peer Assessment 1"
author: "Centrallimit"
output: html_document
keep_md: true
theme: "spacelab"
---


## Loading and preprocessing the data

This document was created on `r Sys.Date()`.

This section describes the data integration process. First of all, 

```{r echo = TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)



setwd("~/Documents/RepData_PeerAssessment1")

df_act <- tbl_df(read.csv("activity.csv", header = TRUE, sep = ","))

mutate(df_act, date=as.Date(date))

```


## What is mean total number of steps taken per day?


```{r echo = TRUE, message=FALSE, warning=FALSE}
steps_sum <- df_act %>% 
        group_by(date) %>%
        summarise(stepssum=sum(steps,na.rm=TRUE))


c_mean <- mean(steps_sum$stepssum, na.rm = TRUE)
c_med <- as.numeric(median(steps_sum$stepssum, na.rm = TRUE))
```
The mean total number of steps taken per day is `r c_mean` and the median is `r c_med`. 


```{r echo = TRUE, message=FALSE, warning=FALSE}
ggplot(steps_sum, aes(x=stepssum)) +
geom_histogram(colour="black", fill="red")
    

```


## What is the average daily activity pattern?

```{r echo = TRUE, message = FALSE, warning = FALSE}
# Grouping and computation of the mean 
steps_day <-  df_act %>%
        group_by(interval) %>%
        summarise(mean_int=mean(steps, na.rm = TRUE))

# Identify the interval with the maximum value of steps
c_max <- steps_day[which.max(steps_day$mean_int), ]    
# Time series plot of daily average 
ggplot(steps_day, aes(interval, mean_int)) + geom_line(color = "red") + xlab("Interval") + ylab("Mean")

```

The interval, that on average across all the days in the dataset, contains the maximum number of steps is `r c_max[1,1]` with an average of `r c_max[1,2]` steps.

## Imputing missing values
First of all, we have to compute the number of missing values. 

```{r echo = TRUE}
# Number of datasets that have missing values
count_c <- sum(!complete.cases(df_act))
# Imputing 
IMP_DF <- df_act %>%
        filter(steps == !complete.cases) %>%
        mutate(steps = c_med)


```
The are `r count_c` datasets with missing values in the sample.


## Are there differences in activity patterns between weekdays and weekends?
